---
title: Portfolio Assignment 2 - Introduction to Cultural Data Science"
output: html_document
authors: Kristiane Uhrenholt Warncke
date: 30/10-2024
---

```{r}
#setup chunk
library(dslabs)
library(ggplot2)
library(dplyr)

```


# Part 1. 

Load  the 'divorce_margarine'dataset  from  the 'dslabs'package.  
```{r}

```


Investigate  the  correlation between  margarine  consumption  and  divorce  rates  in  Maine.  Would  an  increase  in  the preference for margarine lead to skyrocketing divorce rates?

# Part 2.

Load the ‘GSSvocab’dataset from the ‘car’package. This dataset contains people’s scores on an English vocabulary test and includes demographic information.
```{r}

```

Filter for the year 1978and remove rows with missing values (the function na.exclude() is one way to do this – check out the documentation!).
```{r}

```

Is a person’s  score on the vocabulary test  (‘vocab’) significantly impacted by their level  of education (‘educ’)? Visualize the relationship in a plot and build a model. Briefly explain the results. 

```{r}
plot
```

```{r}
model
```

Whether a person is the native of an English-speaking country(‘nativeBorn’) could potentially have an impact on the size of their vocabulary. Visualize the relationship and add the predictor to the model.Briefly explain the results.
```{r}

```



Does  a  person’s  level  of  education  depend  on  whether they  are  a  native  of  the  country? Visualize the relationship. 
```{r}

```


Do you think it makes sense to add the relationship as an interaction term? Try creating the model and briefly explain the results.
```{r}
m1 <- 
m2 <- 
m3 <- 
```


Which model performs best? 
```{r}
anova(m1, m2, m3)
```

When creating plots, try to play around with the options of ggplot2–add titlesand axis labels, and consider using colours.
