---
title: Portfolio Assignment 2 - Introduction to Cultural Data Science"
output: html_document
authors: Kristiane Uhrenholt Warncke
date: 02/11-2024
---

```{r}
#setup chunk
install.packages("dslabs")
install.packages("car")

library(dslabs)
library(ggplot2)
library(dplyr)
library(car)

```


# Part 1. 

**Load  the 'divorce_margarine'dataset  from  the 'dslabs'package.**

```{r}
data("divorce_margarine")
head(divorce_margarine)
```


**Investigate  the  correlation between  margarine  consumption  and  divorce  rates  in  Maine.  Would  an  increase  in  the preference for margarine lead to skyrocketing divorce rates?**

```{r}
ggplot(divorce_margarine,
       aes(x = margarine_consumption_per_capita, y = divorce_rate_maine)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  ggtitle("Margerine Consumption Per Capita as a function of Divorce Rate in Maine") +
  labs(x = "Margerine Consumption Per Capita", y = "Divorce Rate in Maine", caption = "(I didn't know captions were an option! :o)") +
  theme_light()
  

```

```{r}
cor.test(divorce_margarine$margarine_consumption_per_capita, divorce_margarine$divorce_rate_maine)
```

```{r}
summary(lm(divorce_rate_maine ~ margarine_consumption_per_capita, data = divorce_margarine))
```

From the scatter plot visualizing this data set, it seems like there is a positive correlation between the divorce rate in Maine and the margarine consumption per capita. 
The Pearson's correlation test reveals a large positive correlation, where the correlation coefficient is very close to 1 (cor ~ 0.99). Furthermore, from the linear model, it is seen that it is indeed a significant effect (p < 0.001). 
The Estimate suggests that when there is a one unit increase of margarine consumption, the divorce rate in Maine is predicted to increase by ~0.20. 
In terms of the model fit, the R^2 is quite high (R^2 = 0.9852), indicating that this model explains a large portion of the variability in the divorce rate data. 

However, all of these convincing results are likely due to overfitting, as there is not enough data, or a spurious correlation, as there is no logical causal relationship between margarine consumption and divorce rates.


# Part 2.

**Load the ‘GSSvocab’dataset from the ‘car’package. This dataset contains people’s scores on an English vocabulary test and includes demographic information.**
```{r}
data("GSSvocab")
head(GSSvocab)
```

**Filter for the year 1978and remove rows with missing values (the function na.exclude() is one way to do this – check out the documentation!).**

```{r}
dat1978 <- GSSvocab %>%
  filter(year == 1978) %>%
  na.exclude()
```

**Is a person’s  score on the vocabulary test  (‘vocab’) significantly impacted by their level  of education (‘educ’)? Visualize the relationship in a plot and build a model. Briefly explain the results.**

```{r}
ggplot(dat1978, 
       aes(x = educ, y = vocab)) +
  geom_point(position = "jitter") +
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Vocabulary Test Score as a function of Education Level", x = "Education Level", y = "Vocabulary Test Score") +
  theme_light()
```

```{r}
summary(lm(vocab ~ educ, data = dat1978))
```
From the scatter plot it seems that a person’s  score on the vocabulary test increases with level of education. The linear model indicates that for each added level of education, the vocabulary test score increase with 0.39. This positive relation is significant (p < 0.001). However, the linear model does not seem to fit the data very well, (R^2 = 0.2883), indicating that the model explains ~29 % of the variability in the data.


**Whether a person is the native of an English-speaking country(‘nativeBorn’) could potentially have an impact on the size of their vocabulary. Visualize the relationship and add the predictor to the model. Briefly explain the results.**

```{r}
#Visualizing the data
ggplot(dat1978, 
       aes(x = educ, y = vocab, group = nativeBorn)) +
  geom_point(position = "jitter") +
  geom_smooth(method = lm, se = FALSE, aes(color = nativeBorn)) +
  theme_light() +
  labs(title = "Relation between Vocabulary test scores and Education Level by Nativeness", x = "Education Level", y = "Vocabulary Test Score")
```



```{r}
summary(lm(vocab ~ educ + nativeBorn, data = dat1978))
```

From the plot it seems that being a native born in an English-speaking country on average yields a higher vocabulary across all Education Levels. 

The summary of the linear model predicting Vocabulary from both Education Level and Nativeness indicates that whether a person is the native of an English-speaking country or not does in fact have an impact on the size of their vocabulary, and that this effect is significant (p < 0.01). However, this model still only explains ~29 % of the variance in the data.

**Does  a  person’s  level  of  education  depend  on  whether they  are  a  native  of  the  country? Visualize the relationship.** 

```{r}
ggplot(dat1978, 
       aes(x = nativeBorn, y = educ, color = nativeBorn)) +
  geom_point(position = position_jitter(width = 0.4), alpha = 0.5) +
  geom_boxplot(width = 0.3, outlier.shape = NA, color = "black", fill = NA) +
  facet_grid(~ nativeBorn, scales = "free_x", space = "free_x") +
  theme_light() +
  labs(title = "Relation between Education level and Nativeness", x = "Native of an English-speaking country", y = "Education Level")
```
```{r}
ggplot(dat1978, 
       aes(x = educ, fill = nativeBorn)) +
  geom_density(alpha = 0.5) +
  theme_light() +
  labs(title = "Density plot of Education level across Nativeness", x = "Education Level", y = "Density")
```

From the scatter plot with boxplots on top it is seen that there are far more datapoints for Native speakers than non-native speakers. Furthermore, the mean Education Level is similar between the groups, and that there is more variance in the non-native group.

The Density plot similarly reveals that there is a large proportion of the people from Native English-speaking countries that have Education Level 12, and that the non-natives are more spread out across all Education Levels.


**Do you think it makes sense to add the relationship as an interaction term? Try creating the model and briefly explain the results.**
```{r}
summary(lm(vocab ~ educ * nativeBorn, data = dat1978))
```

From the model summary it is seen that:
- Each additional unit of Education Level is associated with an increase of ~0.42 in the vocabulary score, for non Native English-speakers, and this effect is statistically significant (p < 0.001).
- Being a Native English-speaker is in this model associated with an increase of ~0.95 in vocabulary scores, but this effect is not statistically significant (p = 0.186), so we cannot confidently say that being a native English-speaker in itself makes a difference for vocabulary scores when education is considered.
- The interaction between Education Level and Vocabulary score suggests that the effect of Education Level on Vocabulary score is slightly less for native-born individuals compared to non-native-born individuals, with an estimate of ~-0.03. However, this effect is not statistically significant (p = 0.663), meaning that the evidence that the relationship between Education Level and Vocabulary differs based on native-born status is not strong.

To conclude, adding an interaction term does not seem to add significant predictive value for Vocabulary scores, as both the native-born effect and the interaction term are not statistically significant. This suggests that while Education Level positively impacts Vocabulary scores, this effect is consistent across Native and non-Native English-speakers.


**Which model performs best?**
```{r}
m1 <- lm(vocab ~ educ, data = dat1978)
m2 <- lm(vocab ~ educ + nativeBorn, data = dat1978)
m3 <- lm(vocab ~ educ * nativeBorn, data = dat1978)

anova(m1, m2, m3)
```

An ANOVA (analysis of variance) model comparison is conducted with model 1, predicting vocabulary solely from education level, and model 2, having nativeBorn as a predictor as well, and model 3 with an interaction between Education Level and Nativeness. 

The results hereof indicate that adding Nativeness as a predictor significantly increases the models ability to predict Vocabulary test scores. The p-value here describes the probability of observing the data if the additional variable (nativeBorn) had no effect. It is seen that it is a statistically significant effect (p < 0.01), indicating that adding Nativeness improves the model fit. 
However, model 3 with the interaction term is not seen to have a significant impact on the predictability of Vocabulary (p = 0.663), which was also the conclusion above. 

Thus, model 2 performs best out of the three linear models. 

