---
title: An R Markdown document converted from "stats_3_empty_ipynb.ipynb"
output: html_document
---

```{r}
install.packages('dslabs')
```


```{r}
library(ggplot2)
library(dplyr)
library(dslabs)
```


# Exercise 1 - introducing correlations 

Let's look at the relationship between FERTILITY and LIFE EXPECTANCY. For this we'll use the gapminder dataset
```{r}
data(gapminder)
```

```{r}
head(gapminder)
```

```{r}
str(gapminder)
```

Let's restrict our question a bit, and start by looking just at the year 2000
```{r}
dat2000 <- gapminder %>% 
  filter(year==2000)
str(dat2000)
```

Let's start by just looking at the correlation. The best way to start is to visualise the relationship
```{r}
ggplot(dat2000,
       aes(x = fertility, y = life_expectancy)) +
  geom_point()
```

```{r}
cor.test(dat2000$fertility, dat2000$life_expectancy)
```

Let's look at another question. What about the relationship between population and life expectancy?

Let's start by just looking at the correlation. The best way to start is to visualise the relationship
```{r}
ggplot(dat2000, 
       aes(x = population, y = life_expectancy)) +
  geom_point()
```

This plot is hard to read due to the two very high data points, let's censor them
```{r}
ggplot(dat2000, 
       aes(x = population, y = life_expectancy)) +
  geom_point() +
  xlim(0, 2e8)
```

```{r}
cor.test(dat2000$life_expectancy, dat2000$population)
```

Let's apply linear models to both of these relationships
```{r}
fertility_model <- lm(life_expectancy ~ fertility, dat2000)
summary(fertility_model)
```

```{r}
population_model <- lm(life_expectancy ~ population, dat2000)
summary(population_model)
```

# Exercise 2 - interpreting linear models  

Let's go back to last week. We had a question about the influence of number of cylinders on fuel economy. Let's load the data and return to that question

```{r}
install.packages("fueleconomy")
library(fueleconomy)
```

```{r}
data("vehicles")
```

```{r}
head(vehicles)
```

Let's isolate the relevant data again for Audi and VW midsizes
```{r}
Audi_VW <- vehicles %>%
  filter(make == "Audi" & class == "Midsize Cars" | make == "Volkswagen" & class == "Midsize Cars")
str(Audi_VW)
```

```{r}
ggplot(Audi_VW, 
       aes(x = cyl, 
           y = cty)) +
  geom_point(position = 'jitter') +
  facet_wrap( ~ make)
```

Let's use linear models to investigate the effect of number of cylinders on fuel economy in both makes
```{r}
summary(lm(cty ~ cyl, data = subset(Audi_VW, make == "Volkswagen")))
```

```{r}
summary(lm(cty ~ cyl, data = subset(Audi_VW, make == "Audi")))
```

Let's look at this INTERACTION between variables WITHIN THE MODEL
```{r}
summary(lm(cty ~ cyl * make, Audi_VW))
```

Let's go back to this question about which model is better. We can do this using model comparison. Let's build a model without make, and one with, and see what we learn
```{r}
m1 <- lm(cty ~ cyl, Audi_VW)
m2 <- lm(cty ~ cyl + make, Audi_VW)
m3 <- lm(cty ~ cyl * make, Audi_VW)

anova(m1, m2, m3)
```

# Exercise 3 - non-linear predictors

Lets look at the relationship infant mortality and life expectancy in the year 2000 what do you notice?
```{r}
ggplot(dat2000,
       aes(x = infant_mortality, 
           y = life_expectancy)) +
  geom_point()
```

```{r}
dat2000$infant_mortality2 <- dat2000$infant_mortality^2 #making infant mortality an exponential variable 
```

```{r}
str(dat2000) #check that it has been added
```

```{r}
m4 <- lm(life_expectancy ~ infant_mortality, dat2000)
m5 <- lm(life_expectancy ~ infant_mortality + infant_mortality2, dat2000)

anova(m4, m5) #see that the model with the exponential variable explains more variance, and is therefore a better fit. 
```








